---
title: "Homework 2"
author: "Lydia Parr" 
date: 2019-10-09
output: 
  github_document
---

## Setup 

First, I am loading relevant packages. 

```{r setup}
library(tidyverse)
library(readxl)
library(dplyr)
```

## Problem 1

I placed the Mr. Trash Wheel data set into a subdirectory named hw2_data within this one, p8105_hw2_lcp2148. I renamed it trash_wheel.xlsx for ease of use. 

Next, I am reading in the Excel file and cleaning by: 

- specifying the sheet in the Excel file 
- omitting non-data entries (rows with notes / figures; columns containing notes) 
- omitting rows that do not include dumpster-specific data
- converting variable names to snake case
- rounding the number of sports balls to the nearest integer

Note: the row numbers are all off because it used the first column instead of excel columns, I do not have the time or mental strength to fix this and also attempt the rest of the questions.

```{r mr_trash}
mr_trash = read_excel("./hw2_data/trash_wheel.xlsx", sheet = "Mr. Trash Wheel")[-c(1, 11,21,29,35,39,45,50,54,58,60,64,73,80,102,109,116,122,126,128,137,139,146,152,158,162,169,178,184,188,191,195,200,202,205,210,229,238,242,248,257,260,263,266,269,279,282,300,312,323,337,343,352,356,366,371,377,384,391,403,407,420,428,437,445,448,458,460,464,473,477,490,498,508,513,517,532,534,535),-c(15:17)]  %>%
  janitor::clean_names() %>% 
   mutate(sports_balls = round(sports_balls))
``` 

Now I am reading and cleaning precipitation data for 2018 and 2019. I will omit rows without precipitation data and add a variable for year before combining the precipitation data sets and converting month to a character variable.

```{r precip}
precip2018 = read_excel("./hw2_data/trash_wheel.xlsx", sheet = "2018 Precipitation")[-c(14),]  
precip2018 = mutate(precip2018, 
       year = 2018) %>% janitor::clean_names()

precip2019 = read_excel("./hw2_data/trash_wheel.xlsx", sheet = "2019 Precipitation")[-c(14),] 
precip2019 = mutate(precip2019, 
       year = 2019) %>% janitor::clean_names()

precipboth = bind_rows(precip2018, precip2019)[-c(1,14),]

rename(precipboth, month = precipitation_in, inches_precip = x2)
##rename doesn't appear to work in env, probably not even supposed to 
##do it this way, sorry

precipboth$precipitation_in = as.character("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December" )
month.name

``` 

Write a paragraph about these data; you are encouraged to use inline R. 
`summmary(precipboth)` Be sure to note the number of observations in both resulting datasets, and give examples of key variables. For available data, what was the total precipitation in 2018 70.33 in? The median number of sports balls in a dumpster in 2019 `median mr_trash$sports_balls when year in(2019)`


```{r mr_t}



```

## Problem 2

```{r pressure}

```
