---
title: "Homework 2"
author: "Lydia Parr" 
date: 2019-10-09
output: 
  github_document
---

## Setup 

First, I am loading relevant packages. 

```{r setup}
library(tidyverse)
library(readxl)
library(dplyr)
```

## Problem 1

I placed the Mr. Trash Wheel data set into a subdirectory named hw2_data within this one, p8105_hw2_lcp2148. I renamed it trash_wheel.xlsx for ease of use. 

Next, I am reading in the Excel file and cleaning by: 

- specifying the sheet in the Excel file 
- omitting non-data entries (rows with notes / figures; columns containing notes) 
- omitting rows that do not include dumpster-specific data
- converting variable names to snake case
- rounding the number of sports balls to the nearest integer

Note: the row numbers are all off because it used the first column instead of excel columns, I do not have the time or mental strength to fix this and also attempt the rest of the questions.

```{r mr_trash}
mr_trash = read_excel("./hw2_data/trash_wheel.xlsx", sheet = "Mr. Trash Wheel")[-c(1, 11,21,29,35,39,45,50,54,58,60,64,73,80,102,109,116,122,126,128,137,139,146,152,158,162,169,178,184,188,191,195,200,202,205,210,229,238,242,248,257,260,263,266,269,279,282,300,312,323,337,343,352,356,366,371,377,384,391,403,407,420,428,437,445,448,458,460,464,473,477,490,498,508,513,517,532,534,535),-c(15:17)]  %>%
  janitor::clean_names() %>% 
   mutate(sports_balls = round(sports_balls))
``` 

Now I am reading and cleaning precipitation data for 2018 and 2019. I will omit rows without precipitation data and add a variable for year before combining the precipitation data sets and converting month to a character variable.

```{r precip}
precip2018 = read_excel("./hw2_data/trash_wheel.xlsx", sheet = "2018 Precipitation")[-c(14),]  
precip2018 = mutate(precip2018, 
       year = 2018) %>% janitor::clean_names()

precip2019 = read_excel("./hw2_data/trash_wheel.xlsx", sheet = "2019 Precipitation")[-c(14),] 
precip2019 = mutate(precip2019, 
       year = 2019) %>% janitor::clean_names()

precipboth = bind_rows(precip2018, precip2019)[-c(1,14),]

rename(precipboth, month = precipitation_in, inches_precip = x2)
##rename doesn't appear to work in env, probably not even supposed to 
##do it this way, sorry

precipboth$precipitation_in 
month.name

``` 

The data look terrible because I wasn't able to clean any of the data sets properly. 
Key variables in Mr Trash Wheel include weight in tons, volume in cubic years, homes powers, and number of trash items collected such as plastic bottles, cigarette butts, and grocery bags (and of course, sports balls). `summary(mr_trash)` The median weight was 3.380 tons, and the mean was 9.037 tons. `print(mr_trash)` The number of observations in this data set is 456. The median number of sports balls in a dumpster in 2019 `if (mr_trash$year <- 2019) {print median mr_trash$sports_balls}` 

`summary(precipboth)` This data set is the worst because I'm so bad at R and couldn't do anything except put the data together and remove two rows! Yay. The variables should be year, precipitation in inches, and months. Total precipitation in 2018 was `print median precip2018$x2` 70.33 in. 


```{r mr_t}



```

## Problem 2

```{r pressure}

```
