---
title: "Homework 2"
author: "Lydia Parr" 
date: 2019-10-09
output: 
  github_document
---

## Setup 

First, I am loading relevant packages. 

```{r setup}

library(tidyverse)
library(readxl)
library(dplyr)
library(ggplot2)
library(knitr)

```

## Problem 1

I placed the Mr. Trash Wheel data set into a subdirectory named hw2_data
within this one, p8105_hw2_lcp2148. I renamed it `trash_wheel.xlsx` for 
ease of use. 

Next, I am reading in the Excel file and cleaning by: 

- specifying the sheet in the Excel file 
- omitting non-data entries (rows with notes / figures; columns containing
notes) 
- omitting rows that do not include dumpster-specific data
- converting variable names to snake case
- rounding the number of sports balls to the nearest integer

```{r mr_trash}

mr_trash = read_excel("./hw2_data/trash_wheel.xlsx", sheet = "Mr. Trash Wheel", cellranger::cell_cols(1:14)) %>% drop_na() %>%
  janitor::clean_names() %>% 
   mutate(sports_balls = round(sports_balls))

``` 

Now I am reading and cleaning precipitation data for 2018 and 2019. I will
omit rows without precipitation data and add a variable for year before
combining the precipitation data sets and converting month to a
character variable.

```{r precip}

precip2018 = read_excel("./hw2_data/trash_wheel.xlsx", sheet = "2018 Precipitation", skip = 1)
precip2018 = mutate(precip2018, 
       year = 2018) %>% janitor::clean_names()

precip2019 = read_excel("./hw2_data/trash_wheel.xlsx", sheet = "2019 Precipitation", skip = 1)
precip2019 = mutate(precip2019, 
       year = 2019) %>% janitor::clean_names()

precipboth = bind_rows(precip2018, precip2019) %>% drop_na() 

dfmnth = tibble(
  month = 1:12, 
  mnth = c(month.name[month]))

precipfinal = left_join(precipboth, dfmnth, by = "month") %>% 
  select(-c(month)) %>% 
  rename(month = mnth, precip_in = total)

``` 

Key variables in Mr Trash Wheel include `r names(mr_trash)`.  The number of observations in this data set is 456. The median number of sports balls in a dumpster in 2019 `r mr_trash %>% filter(year == 2019) %>% pull(sports_balls) %>% median()`.

For the precipitation data set, the variables are `r names(precipfinal)`. Total precipitation in 2018 was `r precipfinal %>% filter(year == 2018) %>% pull(precip_in) %>% sum()` in. 

## Problem 2

First I am importing pols-month.csv, unemployment.csv, and snp.csv. 
tasks-- month and year will be keys to merge data sets
I cleaned the data in pols-month.csv. Use separate() to break up the variable mon into integer variables year, month, and day; replace month number with month name; create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable.

Second, clean the data in snp.csv using a similar process to the above. For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.

Third, tidy the unemployment data so that it can be merged with the previous datasets. This process will involve switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.

Join the datasets by merging snp into pols, and merging unemployment into the result.
  

```{r fivtheight}

pols_month = read_csv(file = "./hw2_data/pols-month.csv") %>%
  janitor::clean_names() %>% 
  pivot_longer(
    cols = starts_with("prez_"),
    names_prefix = "prez_",
    names_to = "president")  %>%
  mutate(year = lubridate::year(mon), 
                month = lubridate::month(mon), 
                day = lubridate::day(mon)) %>% 
  left_join( dfmnth, by = "month")  %>% 
  select(-c(day, month))

  
unempl = read_csv(file = "./hw2_data/unemployment.csv") %>%
    janitor::clean_names() %>%
  pivot_longer(
    jan:dec,
    names_to = "mnth",
    values_to = "unemployment") %>%
  mutate(mnth = recode(mnth,
                       jan = "January", feb = "February", mar = "March", 
                       apr = "April", may = "May", jun = "June", 
                       july = "July", aug = "August", sep = "September", 
                       oct = "October", nov = "November", dec = "December"))
  
snp = read_csv(file = "./hw2_data/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, into = c("month", "day", "year"), convert = TRUE) %>%
  mutate(year = ifelse(year > 15, year + 1900, year + 2000)) %>% 
  left_join(dfmnth, by = "month")  %>% 
  select(-c(day, month))

merged_538 =
  left_join(snp, pols_month, by = c("year", "mnth"))  %>% 
  left_join(., unempl, by = c("year", "mnth")) %>% 
  rename(month = mnth) %>% drop_na() %>% 
  select(-c(mon))

```

The `pols_month` data set contains data on the party affiliation of the current president and number of republican and democratic governors, senators, and senators on specific dates, which were changed to indicate year and month. The representative of the merged data set, `merged_538`, is `r dim(merged_538)`. The range of years in this merged data set is `r merged_538 %>% select(year) %>% range`. The names of key variables are month and year, out of the total variables in the data set, `r merged_538 %>% names`. 

## Problem 3 

Next, I'm importing the baby names data and cleaning by correcting ethnicity names typos to match, making the case structure of the name variable consistent, and removing the duplicate rows. 

```{r baby_names}

baby_names = read_csv(file = "./hw2_data/Popular_Baby_Names.csv", col_types = "dcccdd")  %>%
  janitor::clean_names() %>%
  distinct() %>%
  mutate(childs_first_name = tolower(childs_first_name)) %>%
  mutate(gender = tolower(gender)) %>%
  mutate(ethnicity = tolower(ethnicity)) 

baby_names$ethnicity[baby_names$ethnicity == "asian and paci"] = "asian and pacific islander" 
baby_names$ethnicity[baby_names$ethnicity == "black non hisp"] = "black non hispanic" 
baby_names$ethnicity[baby_names$ethnicity == "white non hisp"] = "white non hispanic"
  
```

Next, I make a table using `pivot_wider` showing the rank in popularity of the 
name “Olivia” as a female baby name over time including rows for ethnicities 
and columns for year. 

I then make a table using a similar process showing the most popular name among 
male children over time.

For male, white non-hispanic children born in 2016, I use `ggplot` to make a 
scatter plot showing the number of children with a name against the rank in 
popularity of that name.
## %>%
##knitr::kable(format = "pipe", col.names = gsub("[_]", " ", names, caption = "Popularity of  ##the name Olivia over time")
 

```{r baby_figures}

baby_names %>%  
  select(childs_first_name, gender, ethnicity, year_of_birth, rank) %>%  
  filter(childs_first_name == "olivia", gender == "female")  %>%
  arrange(year_of_birth)  %>%
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
  ) 

baby_names %>%  
  select(childs_first_name, gender, ethnicity, year_of_birth, rank) %>%  
  filter(gender == "male", rank == 1)  %>%
  arrange(year_of_birth)  %>%
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name"
  ) 

baby_names %>% filter(ethnicity == "white non hispanic", year_of_birth == 2016, 
                      gender == "male")  %>% ggplot(aes(rank, count)) + geom_point(size = 0.5) + ggtitle("Number and rank of names among white male children born in 2016")

```
